---
title: "Spam Email" 
author: "Amy Mitchell-Whittington"
date: 2023-08-15
categories: ["Tidy Tuesday"]
tags: ["TidyTuesday"]
image:
  caption: ''
  focal_point: ''
  preview_only: no
summary: xx
---

## Spam Email [</>](https://github.com/amymitchell-whittington/TidyTuesday/blob/main/2023/08-08-2023-hotones/08-08-2023-hotones.R)  

The data this week comes from Vincent Arel-Bundock's [Rdatasets package] (https://vincentarelbundock.github.io/Rdatasets/index.html).

> *Rdatasets is a collection of 2246 datasets which were originally distributed alongside the statistical software
> environment R and some of its add-on packages. The goal is to make these data more broadly accessible for teaching
> *and statistical software development.

We're working with the spam email dataset. This is a subset of the spam e-mail database.

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 


```

```{r explore, echo=FALSE, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 



```

```{r plot1, echo=FALSE, include = TRUE, message=FALSE, dpi = 300}



```
