---
title: "U.S. Populated Places" 
author: "Amy Mitchell-Whittington"
date: 2023-06-27
categories: ["Tidy Tuesday"]
tags: ["TidyTuesday"]
image:
  caption: ''
  focal_point: ''
  preview_only: no
summary: A map of all the U.S. counties / landmarks our friends are visiting for their honeymoon.
---

## U.S.Populated Places [</>](https://github.com/amymitchell-whittington/TidyTuesday/blob/main/2023/27-06-2023-USpopulatedplaces/27-06-2023USpopulatedplaces.Rmd)  

Data this week comes from the [National Map Staged Products Directory](https://prd-tnm.s3.amazonaws.com/index.html?prefix=StagedProducts/GeographicNames/) from the [US Board of Geographic Names.](https://www.usgs.gov/us-board-on-geographic-names/download-gnis-data). 

This was a fun #TidyTuesday data set to work with because it timed well with some friends who flew to the U.S. last week to have an Elvis wedding in Las Vegas : heart :

They're hiring an RV and hitting the road for a few weeks, with plans to stop at a few major national parks in the area, so I thought I would use their honeymoon travel itinerary as a guide to visualising some of this week's data. 

Using the usmap package was a little tricky, as I had to transform my data so I could plot it properly. I think after my attempt at using a different maps package to plot reported UFO sightings onto a map of Australia last week, I might need to try out ggmaps to see if it's a little more streamlined.

I ran out of time to try and tweak the plot .png, as it is a lot narrower than my previous knitted plots. I'll have to do some research into why that is and try to remedy it for next week's data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

library(ggplot2)
library(tidyr)
library(ggtext)
library(lubridate)
library(dplyr)
library(readr)
library(maps)
library(mapdata)
library(ggrepel)
library(usmap)

us_place_names <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-06-27/us_place_names.csv')
us_place_history <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-06-27/us_place_history.csv')

data <- us_place_names%>%
  left_join(us_place_history)

```

```{r explore, echo=FALSE, dpi=300}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

roadtrip <- data %>%
  filter(state_name %in% c("Arizona", "California", "Utah", "Nevada", "Oregon"),
         history != "NA",
         feature_name %in% c("Las Vegas", "Death Valley Junction", "South Lake Tahoe", "Portland", "Redwood", "Big Sur", "Los Angeles", "Grand Canyon", "Bryce Canyon City")) 

roadtrip_lonlat <- roadtrip %>%
  select(prim_lat_dec, prim_long_dec, feature_name, county_name) %>%
  rename(lat = prim_lat_dec, lon = prim_long_dec) %>%
    unite("label", county_name:feature_name, sep = " / ")

label <- roadtrip_lonlat$label
  
```

```{r plot, echo=FALSE, include = TRUE, message=FALSE, dpi = 300}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

transformed_data <- usmap_transform(roadtrip_lonlat)

plot_usmap(include = c("CA", "AZ", "UT", "NV", "OR"),
             labels = TRUE, label_color = "cornsilk",
                  fill = "cornsilk", alpha = 0.25, color = "darkcyan", linewidth = 1) + 
  ggrepel::geom_label_repel(data = transformed_data,
             aes(x = x, y = y, label = label),
             size = 2, alpha = 0.8, 
             fill = "cornsilk",
             label.size = 0.5) +
   geom_point(data = transformed_data,
             aes(x = x, y = y),
             color = "purple", alpha = 2) +
    labs(title = "The RV honeymoon circuit",
         subtitle = "A map of all the U.S. counties / landmarks our friends\nare visiting in an RV for their honeymoon.",
      caption = "Source: US Board of Geographic Names.") +
  theme(plot.title = element_markdown(size = 13, colour = "cornsilk", face = "bold"),
        plot.subtitle = element_text(size = 6, colour = "cornsilk"),
        plot.caption = element_text(size = 6, colour = "cornsilk", face = "italic"),
        panel.background = element_rect(fill = "darkcyan"),
        plot.background = element_rect(fill = "darkcyan"))
```

