---
title: "U.S. Historical Markers" 
author: "Amy Mitchell-Whittington"
date: 2023-06-27
categories: ["Tidy Tuesday"]
tags: ["TidyTuesday"]
image:
  caption: ''
  focal_point: ''
  preview_only: no
summary: Tracking missing historical markers in the U.S.
---

## U.S.Historical Markers [</>](https://github.com/amymitchell-whittington/TidyTuesday/blob/main/2023/04-07-2023-historicmarkers/04-07-2023-historicalmarkers.Rmd)  

The data this week comes from the [Historical Marker Database USA Index.](https://www.hmdb.org/geolists.asp?c=United%20States%20of%20America)

This searchable online catalogue was so fun to explore! I didn't realise there was such a thing online (probably because there are only 36 Australian markers recorded on the site compared to 183k U.S. markers).

According to the database, it is *"an illustrated searchable online catalog of historical information viewed through the filter of roadside and other permanent outdoor markers, monuments, and plaques".* Anyone can contribute, either by adding new markers or updating existing marker pages. 

For this Tidy Tuesday project, only the U.S. marker database was provided and what caught my eye the most was the column relating to missing markers. 

I wanted to use percentages because I felt it was the best way to represent the data, and was surprised by the amount of markers that were reported or confirmed missing. I wonder what happened to them?

This was also a great way to test out the labs() annotate function and play around with positioning.


```{r set up, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

library(ggplot2)
library(tidyverse)
library(tidyr)
library(ggtext)
library(lubridate)
library(dplyr)
library(readr)
library(ggrepel)

historical_markers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-07-04/historical_markers.csv')
  
```

```{r explore, echo=FALSE, include = TRUE, message=FALSE, dpi = 300}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

total_markers <- historical_markers %>%
  count(state_or_prov) %>%
  rename(total = n)

missing_markers <- historical_markers %>%
  select(state_or_prov, missing) %>%
  filter(missing != "NA") %>%
  count(state_or_prov) %>%
  rename(missing = n)
  
# % of markers missing (reported and confirmed) per state

perc_markers <- total_markers %>%
  left_join(missing_markers) %>%
  filter(missing != "NA") %>%
  mutate(percentage = (missing/total)*100) %>%
  arrange(desc(percentage)) %>%
  mutate(perc = round(percentage, 1))

perc_markers$label <- paste0("  ",perc_markers$perc,"%")

```

```{r plot, echo=FALSE, message=FALSE, dpi = 300}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

ggplot(data = perc_markers,
       aes(x = reorder(state_or_prov, perc), y = perc)) + #if you add a minus (-) sign to perc in reorder(), you will reverse the order.
  geom_point(colour = "coral4") +
  geom_segment(aes(x = state_or_prov, xend = state_or_prov, y = 0, yend = perc), colour = "cornflowerblue") + 
    geom_text(aes(label = label), hjust = 0, size = 2, colour = "coral4") +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Missing historical markers in the U.S.",
    subtitle = "The percentage of historical markers per state that have been reported or confirmed missing\nvia The Historical Marker Database.",
    caption = "Source: The Historical Marker Database USA Index.",
  ) +
  annotate("text", x = 16, y = 7, label = "Historical markers in the U.S can be cast in metal,\ncarved on stone, or embedded in resin and often mark\nthe location of an historical event or place.", size = 3, colour = "coral4", hjust = 0) +
  theme(
    plot.title = element_text(size = 13, colour = "coral4", face = "bold"),
    plot.subtitle = element_text(size = 9, colour = "coral4"),
    plot.caption = element_text(size = 6, colour = "coral4", face = "italic"),
    plot.background = element_rect(fill = "beige"),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_text(size = 7, colour = "coral4")
  )

```

